---
title: "Homework 1"
author: "Melissa Paniagua"
date: "9/11/2020"
output: html_document
---

This assignment will concentrate on using R and Git.

Let's use data from a finantial entity utilized to perform strategies in a marketing campaing. This dataset has been retrieved from Kaggle.com, and has 17 columns and 11,162 rows.

## Getting data in R

```{r}

#Load the library package
library()

#Load the dataset as a csv file
df<- read.csv("/Users/melissa/Documents/Kent State/Fall 2020/Fund. Machine Learning/datasets-4471-6849-bank.csv")

#Select the first five rows and 10 columns of the dataframe
small.df = df [c(1:5),c(1:10)]

#Summary of the small data frame
print (small.df)

#See the data frame structure
str(df)
```


## Descriptive Statistics

### 1 Quantitative variables ###

**1.1 Customers' age**

```{r}
#Identify type of object
class(df$age)
#Statistics summary
summary(df$age)
#Draw a histogram to analyze the customers' age pattern
hist(df$age, 
     main="Histogram for Ages", 
     xlab="Age", 
     border="darkblue", 
     col="lightblue",
     xlim=c(15,95),
     ylab="Frequency of Ages",
     las=1, 
     breaks=20)
```

Here, we can identify that most customers are between 25 and 60 years old. This histogram has a right tail, which means a few customers are elderly people, but the majority are young.

**1.2 Customers' balance**

```{r}
#Identify type of object
class(df$balance)
#Statistics summary
summary(df$balance)
#Draw a histogram to analyze the customers' balance pattern.
hist(df$balance, 
     main="Histogram of Individual's Balance", 
     xlab="Balance", 
     border="darkgreen", 
     col="lightgreen",
     xlim=c(-7000,82000),
     ylab="Frequency of Balances",
     las=1, 
     breaks=10)
```

Here, we can identify that most customers balance is between 0 to 1000. This histogram has a right tail, which means that a few customers how has a great balance around 8000, but those are outliers. Something to take into consideration is that there are a considerable number of customers with negative balance.

### 2 Categorical variables

**2.1 Customers' marital status**
```{r}
#Identify type of object
class(df$marital)
#Summary of marital status
table(df$marital)
#Pie chart of marital status
pie(table(df$marital), main= "Pie Chart of Marital Status")

```

More than 50% of the customers are married.

**2.2 Customers' education**
```{r}
#Identify type of object
class(df$education)
#Summary of education
table(df$education)
#Bar chart of education
barplot(table(df$education),
        main= "Bar Chart of education",
        col= c("darkblue"))
```

The graph show above, confirms that % customers have 

## Variable Transformation
```{r}
#Current data type
class(df$education)
#Data type transformation from factor to character. Show only the first 6 rows.
head(as.character(df$job))
#Current data type
class(as.character(df$job))
```


## Visualizations

### Plot
```{r}
# Housing vs Education"
counts <- table(df$housing, df$education)
barplot(counts, main="Housing vs Education",
        xlab="Education", 
        col=c("darkred","green"),
        legend = rownames(counts), beside=TRUE)
```

Here, the bank could identify that most of its customer have secondary and tertiary education.

### Scatterplot
```{r}
#Function to generate a continuous color palette
mycolor <- colorRampPalette(c('red','green'))

#Add the color palette based on customer's balance
mycolor2 <- mycolor(10)[as.numeric(cut(df$balance,breaks = 10))]

#Plot personal Balance vs Age
plot(df$age,df$balance, main="Personal Balance vs Age", 
     xlab="Age", 
     ylab="Balance", 
     col = mycolor2,
     pch = 20)
```

This scatterplot shows the distribution of personal balance and ages. An interesting point, is that the higest balance is a person older than 80 years, which is an outlier in this dataset.



#### Dataset source:
[Moro et al., 2014] S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014

